---
title: "Bayesian Causal Forests (BCF) for Heterogeneous Treatment Effects"
output: rmarkdown::html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(dplyr)
library(ggplot2)
# Optional: install.packages("bcf")
# library(bcf)
```

## 🌲 Bayesian Causal Forests (BCF)

Bayesian Causal Forests (BCF) are a nonparametric Bayesian approach to estimating **individual-level heterogeneous treatment effects** by modeling both the outcome and treatment assignment using tree-based ensembles.

---

## 1. 🧪 Simulate Data with Heterogeneous Effects

```{r}
set.seed(123)
n <- 500

X1 <- rnorm(n)
X2 <- rnorm(n)
X3 <- rbinom(n, 1, 0.5)
tau <- 2 + X1 - 0.5 * X2  # treatment effect varies by covariates
W <- rbinom(n, 1, prob = plogis(0.4 * X1 - 0.3 * X2 + 0.5 * X3))
Y <- tau * W + 1.5 * X1 + 0.8 * X2 + 1.0 * X3 + rnorm(n)

df <- data.frame(X1, X2, X3, W, Y)
```

---

## 2. 📈 Visualize True Heterogeneous Effect

```{r}
ggplot(data.frame(X1, tau), aes(x = X1, y = tau)) +
  geom_point(alpha = 0.4) +
  geom_smooth(se = FALSE) +
  labs(title = "True Heterogeneous Treatment Effect by X1",
       x = "X1", y = "Treatment Effect (tau)") +
  theme_minimal()
```

---

## 3. 📦 Fit a BCF Model

If using the `bcf` package:

```{r eval=FALSE}
X <- as.matrix(df[, c("X1", "X2", "X3")])
bcf_fit <- bcf(y = df$Y, z = df$W, x_control = X, x_moderate = X)
tau_hat <- colMeans(bcf_fit$tau)
```

---

## 4. 🧪 Compare Estimated vs. True Effects

```{r eval=FALSE}
ggplot(data.frame(True = tau, Estimated = tau_hat), aes(x = True, y = Estimated)) +
  geom_point(alpha = 0.4) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "red") +
  labs(title = "True vs. Estimated Treatment Effects",
       x = "True Effect", y = "Estimated Effect") +
  theme_minimal()
```

---

## ✅ Summary

Bayesian Causal Forests:
- Estimate heterogeneous effects without strong parametric assumptions
- Use **Bayesian Additive Regression Trees (BART)** internally
- Require installation of the `bcf` package

---

## 📖 References

- Hahn, Murray, and Carvalho (2020). *Bayesian regression tree models for causal inference: The BCF approach*
- Green & Kern (2012). *Modeling heterogeneous treatment effects with BART*