---
title: "Causal Impact Analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Causal Impact Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(CausalImpact)
library(ggplot2)
```

# 📊 Causal Impact Analysis

Causal Impact is a Bayesian structural time-series method used to evaluate the effect of an intervention when randomization is not feasible.

---

## 🧪 Simulate Pre/Post Intervention Time Series

```{r}
set.seed(101)
n <- 100
pre_period <- 1:70
post_period <- 71:100

x <- arima.sim(model = list(order = c(1, 0, 0), ar = 0.8), n = n)
y <- x + rnorm(n, 0, 0.5)

# Add treatment effect in post period
y[post_period] <- y[post_period] + 2

data_ts <- cbind(y, x)
```

---

## 📈 Run Causal Impact

```{r}
impact <- CausalImpact(data_ts, pre.period = c(1, 70), post.period = c(71, 100))
summary(impact)
plot(impact)
```

---

## ✅ Interpretation

- **CausalImpact** estimates the counterfactual (what would’ve happened without intervention).
- The difference between observed and predicted is the estimated effect.

---

## 📚 References

- Brodersen et al. (2015). *Inferring causal impact using Bayesian structural time-series models*. Annals of Applied Statistics.