---
title: "Causal Impact Analysis for Time Series Interventions"
output: rmarkdown::html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(CausalImpact)
```

## 📈 Causal Impact for Time Series

Causal Impact models estimate the **effect of an intervention** in a time series when there is **no control group**, using Bayesian structural time series models.

---

## 1. 📊 Simulate Time Series Data

```{r}
set.seed(123)
pre_period <- 1:70
post_period <- 71:100
x <- c(rnorm(70, 10, 2), rnorm(30, 15, 2))  # treated series with effect in post period
y <- c(rnorm(70, 10, 2), rnorm(30, 10, 2))  # control series

ts_data <- cbind(y, x)
```

---

## 2. 🧮 Apply Causal Impact

```{r}
impact <- CausalImpact(ts_data, pre_period = c(1, 70), post_period = c(71, 100))
summary(impact)
plot(impact)
```

---

## ✅ Interpretation

- Posterior intervals give uncertainty of effect
- Used for **policy**, **marketing**, and **event evaluation**
- Useful when randomized controls are not available

---

## 📖 References

- Brodersen et al. (2015). *Inferring causal impact using Bayesian structural time-series models*