---
title: "Causal Discovery with DAGitty and Simulated Data"
output: rmarkdown::html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(dagitty)
library(ggdag)
library(dplyr)
library(ggplot2)
```

## 🧭 Causal Discovery: Learning Structure from Data

Causal discovery aims to **infer the causal structure** (i.e. DAG) from observational data. This is useful when domain knowledge is limited or exploratory analysis is needed.

---

## 1. ✨ Simulate Data with Known DAG

```{r}
set.seed(123)
n <- 1000

X <- rnorm(n)
Z <- 0.6 * X + rnorm(n)
Y <- 0.5 * Z + 0.3 * X + rnorm(n)

df <- data.frame(X, Z, Y)
head(df)
```

---

## 2. 🧠 Define the True DAG

```{r}
true_dag <- dagitty("dag {
  X -> Z -> Y
  X -> Y
}")

ggdag(true_dag, layout = "circle") +
  ggtitle("True DAG Used to Generate Data")
```

---

## 3. 🔍 Use `localTests()` to Test Dependencies

We use `dagitty::localTests()` to test conditional independence statements implied by a candidate DAG:

```{r}
test_result <- localTests(true_dag, data = df, type = "cis")
test_result
```

Each row checks whether a conditional independence assumption holds (p > 0.05 supports the claim).

---

## 4. 🧪 Try an Incorrect DAG

```{r}
wrong_dag <- dagitty("dag {
  Z -> X
  Z -> Y
  X -> Y
}")

ggdag(wrong_dag, layout = "circle") +
  ggtitle("Incorrect DAG Assumption")
```

Now test:

```{r}
localTests(wrong_dag, data = df, type = "cis")
```

Look for failed assumptions (very low p-values).

---

## 🧠 Assumptions for Causal Discovery

- **Causal Sufficiency**: All common causes are measured
- **Faithfulness**: Statistical independence reflects causal structure
- **No Measurement Error**

---

## ✅ Summary

Causal discovery lets us:
- Explore structures with little prior knowledge
- Validate assumptions
- Compare competing causal hypotheses

`dagitty` provides a simple entry point. For algorithmic discovery, explore `pcalg`, `bnlearn`, or `cdcs`.

---

## 📖 References

- Textor et al. (2016). *dagitty: Graphical Analysis of Structural Causal Models*
- Spirtes, Glymour, Scheines (2000). *Causation, Prediction, and Search*