---
title: "Covariate Balance Diagnostics in Causal Inference"
output: rmarkdown::html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(dplyr)
library(ggplot2)
library(MatchIt)
library(cobalt)
```

## ⚖️ Covariate Balance Diagnostics

Ensuring that covariates are balanced across treatment groups is critical for reducing bias in causal inference. Diagnostics help verify whether matching or weighting strategies have achieved this.

---

## 1. 🧪 Simulate Confounded Data

```{r}
set.seed(123)
n <- 1000

X1 <- rnorm(n)
X2 <- rbinom(n, 1, 0.5)
logit_p <- -0.5 + 0.7 * X1 + 1.2 * X2
p_treat <- 1 / (1 + exp(-logit_p))
W <- rbinom(n, 1, p_treat)

Y <- 3 * W + 0.5 * X1 + 0.8 * X2 + rnorm(n)
data <- data.frame(X1, X2, W, Y)
```

---

## 2. 🎯 Perform Propensity Score Matching

```{r}
match_model <- matchit(W ~ X1 + X2, data = data, method = "nearest", distance = "logit")
summary(match_model)
```

---

## 3. 📊 Balance Diagnostics with `cobalt`

```{r}
love.plot(match_model, binary = "std", var.order = "unadjusted",
          abs = TRUE, thresholds = c(m = .1),
          title = "Covariate Balance Before and After Matching")
```

---

## 4. ✅ Interpreting Balance Metrics

- **Standardized Mean Differences (SMD)**: < 0.1 considered balanced
- Look for reduction in SMD after matching
- Avoid overfitting and extreme weights

---

## 📖 References

- Austin (2009). *Balance diagnostics for comparing the distribution of baseline covariates using standardized differences*
- Stuart (2010). *Matching Methods for Causal Inference: A Review and a Look Forward*