---
title: "DAGs and Assumptions in Causal Inference"
output: rmarkdown::html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(dagitty)
library(ggdag)
library(ggplot2)
```

## 🔗 DAGs and Assumptions in Causal Inference

Directed Acyclic Graphs (DAGs) provide a powerful way to encode assumptions about how variables in your system are related.

They help us:
- Identify confounders and mediators
- Test identification strategies
- Derive adjustment sets

---

## 🎯 Define a Simple DAG

Let’s say we're estimating the effect of `Treatment` on `Outcome`, with a confounder `X`.

```{r}
dag <- dagitty("dag {
  X -> Treatment
  X -> Outcome
  Treatment -> Outcome
}")

plot(dag)
```

---

## 📚 Adjustment Sets

Which variables do we need to control for to identify the causal effect of `Treatment` on `Outcome`?

```{r}
adjustmentSets(dag, exposure = "Treatment", outcome = "Outcome")
```

This shows that adjusting for `X` is sufficient.

---

## 🧠 More Complex DAG

```{r}
complex_dag <- dagitty("dag {
  Age -> Treatment
  Income -> Treatment
  Age -> Outcome
  Income -> Outcome
  Treatment -> Behavior -> Outcome
  unobserved1 [unobserved]
  unobserved1 -> Income
  unobserved1 -> Outcome
}")

ggdag::ggdag(complex_dag, layout = "circle") +
  ggtitle("DAG with Confounding and Mediation")
```

---

## ✅ Summary

DAGs help us:
- Encode our **causal assumptions**
- Identify valid **adjustment strategies**
- Prevent common pitfalls like **conditioning on colliders**

Learning to draw and analyze DAGs is a core skill for any causal analyst.

---

## 📖 References

- Textor et al. (2016). *dagitty: Graphical Analysis of Structural Causal Models*
- Pearl (2009). *Causality: Models, Reasoning, and Inference*