---
title: "Difference-in-Differences (DiD): Simulated Example"
output: rmarkdown::html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(dplyr)
library(ggplot2)
```

## 📊 Difference-in-Differences (DiD)

DiD is a popular method to estimate causal effects using **panel or repeated cross-sectional data** before and after a treatment is applied to one group but not another.

---

## 1. ✨ Simulate DiD Setup

```{r}
set.seed(123)
n <- 1000
group <- rep(c("control", "treatment"), each = n/2)
time <- rep(c("pre", "post"), times = n/2)
treated <- ifelse(group == "treatment" & time == "post", 1, 0)

baseline <- rnorm(n)
Y <- 2 * treated + 0.5 * as.numeric(group == "treatment") + 
     0.3 * as.numeric(time == "post") + baseline + rnorm(n, 0, 0.5)

df <- data.frame(group, time, treated, Y)
head(df)
```

---

## 2. 📈 Visualize Group-Time Means

```{r}
df_summary <- df %>%
  group_by(group, time) %>%
  summarise(mean_Y = mean(Y), .groups = "drop")

ggplot(df_summary, aes(x = time, y = mean_Y, group = group, color = group)) +
  geom_line(size = 1.2) +
  geom_point(size = 2) +
  labs(title = "Group-Time Averages", y = "Mean Outcome", x = "Time") +
  theme_minimal()
```

---

## 3. 🧮 Estimate DiD Effect

```{r}
df$time_num <- ifelse(df$time == "post", 1, 0)
df$group_num <- ifelse(df$group == "treatment", 1, 0)

did_model <- lm(Y ~ group_num * time_num, data = df)
summary(did_model)
```

The coefficient on `group_num:time_num` is the **DiD estimate**.

---

## ✅ Assumptions

- **Parallel Trends**: Control and treatment groups would evolve similarly in absence of treatment
- **No Spillover Effects**
- **No Anticipation**

---

## 📖 References

- Angrist & Pischke (2009). *Mostly Harmless Econometrics*
- Bertrand, Duflo, Mullainathan (2004). *How Much Should We Trust Differences-in-Differences Estimates?*