---
title: "Double Machine Learning (DML) for Causal Inference"
output: rmarkdown::html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(dplyr)
library(ggplot2)
library(glmnet)
library(caret)
```

## ðŸ¤– Double Machine Learning (DML)

Double Machine Learning (DML) uses flexible machine learning models to estimate nuisance parameters (e.g., propensity scores, outcome models) and removes their bias via **orthogonalization**, enabling valid causal inference.

---

## 1. ðŸ§ª Simulate Data with Confounding

```{r}
set.seed(123)
n <- 1000

X1 <- rnorm(n)
X2 <- rnorm(n)
W <- rbinom(n, 1, prob = plogis(0.5 * X1 + 0.5 * X2))
Y <- 2 * W + 1.5 * X1 + 1.0 * X2 + rnorm(n)

df <- data.frame(X1, X2, W, Y)
```

---

## 2. ðŸ” DML Steps

### (a) Split into folds

```{r}
folds <- createFolds(df$Y, k = 2)
train1 <- df[folds[[1]], ]
train2 <- df[folds[[2]], ]
```

### (b) Estimate nuisance functions

```{r}
# Outcome model: Y ~ X1 + X2 (train2 on train1)
y_model <- cv.glmnet(as.matrix(train1[, c("X1", "X2")]), train1$Y)
m_Y2 <- predict(y_model, as.matrix(train2[, c("X1", "X2")]), s = "lambda.min")

# Treatment model: W ~ X1 + X2
w_model <- cv.glmnet(as.matrix(train1[, c("X1", "X2")]), train1$W, family = "binomial")
m_W2 <- predict(w_model, as.matrix(train2[, c("X1", "X2")]), s = "lambda.min", type = "response")
```

### (c) Orthogonalization

```{r}
res_Y <- train2$Y - m_Y2
res_W <- train2$W - m_W2

alpha_hat <- coef(lm(res_Y ~ res_W))[2]
cat("Estimated Treatment Effect (DML):", round(alpha_hat, 3))
```

---

## âœ… Summary

Double Machine Learning:
- Uses ML to estimate nuisance models (Y|X and W|X)
- Debiases via residual-on-residual regression
- Yields valid causal estimates with flexible modeling

---

## ðŸ“– References

- Chernozhukov et al. (2018). *Double/Debiased Machine Learning for Treatment and Structural Parameters*
- Athey & Imbens (2019). *Machine Learning and Econometrics*