---
title: "Heterogeneous Treatment Effects"
output: rmarkdown::html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(CausalInvestData)
library(dplyr)
library(ggplot2)
```

## 🧪 What Are Heterogeneous Treatment Effects?

Heterogeneous Treatment Effects (HTEs) occur when the impact of a treatment varies across different units or subgroups in the population. In financial decision-making, understanding **who benefits more (or less)** from a given strategy can be more useful than knowing the average effect.

We’ll explore this using the `fund_performance` dataset.

---

## 📦 Load the Data

```{r}
data("fund_performance")
head(fund_performance)
```

Let’s inspect how treatment effects differ by levels of `beta` (a proxy for fund risk exposure):

---

## 📊 Stratify by Fund Beta Quantiles

```{r}
fund_performance <- fund_performance %>%
  mutate(beta_group = ntile(beta, 3))  # low, medium, high

group_summary <- fund_performance %>%
  group_by(beta_group, treatment) %>%
  summarise(mean_return = mean(return), .groups = "drop")

group_summary %>%
  tidyr::pivot_wider(names_from = treatment, values_from = mean_return,
                     names_prefix = "treatment_") %>%
  mutate(estimated_effect = treatment_1 - treatment_0)
```

---

## 📈 Visualize Treatment Effects by Risk Exposure

```{r}
group_summary %>%
  ggplot(aes(x = factor(beta_group), y = mean_return, fill = factor(treatment)))
  geom_col(position = "dodge")
  labs(title = "Returns by Treatment Across Beta Groups",
       x = "Beta Quantile Group", y = "Mean Return",
       fill = "Treatment")
  theme_minimal()
```

---

## 🌲 Optional: Use Causal Forests (`grf`)

If you have the `grf` package installed, you can run:

```r
# Uncomment and install if needed
# install.packages("grf")

# library(grf)
# X <- fund_performance %>% select(market_return, alpha, beta)
# Y <- fund_performance$return
# W <- fund_performance$treatment
# forest <- causal_forest(X, Y, W)
# tau_hat <- predict(forest)$predictions
# hist(tau_hat, breaks = 30, main = "Estimated Treatment Effects (GRF)", xlab = "tau_hat")
```

---

## ✅ Summary

| Subgroup (by beta) | Estimated Effect |
|--------------------|------------------|
| Low beta funds     | ?                |
| Mid beta funds     | ?                |
| High beta funds    | ?                |

Even without advanced ML, subgroup analysis reveals **important insights** on how treatments work differently for different fund profiles.

---

## 📖 Citation

```{r}
citation("CausalInvestData")
```