---
title: "Introduction to CausalInvestData"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to CausalInvestData}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(CausalInvestData)
```

## Overview

`CausalInvestData` provides simulated datasets for causal inference in institutional investment management. The package includes four core datasets designed to reflect real-world structures, enabling users to prototype, teach, and evaluate methods such as propensity score matching, causal forests, and impact analysis.

---

## Dataset: `fund_performance`

This dataset simulates returns for 1,000 investment funds, with treatment variables for active strategy testing.

```{r}
data("fund_performance", package = "CausalInvestData")
head(fund_performance)
```

### Propensity Score Matching Example

```{r}
library(MatchIt)
m.out <- matchit(treatment ~ market_return + alpha + beta, data = fund_performance)
summary(m.out)
```

---

## Dataset: `portfolio_allocations`

This dataset simulates allocation profiles across 500 portfolios with varying risk levels and strategic treatments.

```{r}
data("portfolio_allocations", package = "CausalInvestData")
head(portfolio_allocations)
```

---

## Dataset: `client_behavior`

This dataset simulates client demographics, satisfaction scores, income, and whether they churned, with treatment for receiving financial advice.

```{r}
data("client_behavior", package = "CausalInvestData")
head(client_behavior)
```

---

## Dataset: `macro_shocks`

Monthly macroeconomic data (interest rates, GDP growth, market index) to model the effect of macro shocks on financial strategies.

```{r}
data("macro_shocks", package = "CausalInvestData")
head(macro_shocks)
```

---

## Summary

This package is ideal for:

- Financial data scientists building causal ML pipelines
- Academics teaching causal inference methods
- Practitioners evaluating financial interventions

To cite the package, run:

```{r}
citation("CausalInvestData")
```
