---
title: "Propensity Score Weighting (PSW) for Causal Inference"
output: rmarkdown::html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(dplyr)
library(ggplot2)
library(survey)
```

## ⚖️ Propensity Score Weighting (PSW)

PSW is a technique used to adjust for confounding in observational studies by **reweighting units** based on their estimated probability of receiving treatment.

---

## 1. 🧪 Simulate Confounded Data

```{r}
set.seed(123)
n <- 1000

X1 <- rnorm(n)
X2 <- rbinom(n, 1, 0.5)
logit_p <- -0.5 + 0.7 * X1 + 1.2 * X2
p_treat <- 1 / (1 + exp(-logit_p))
W <- rbinom(n, 1, p_treat)

# True treatment effect is 3
Y <- 3 * W + 0.5 * X1 + 0.8 * X2 + rnorm(n)

data <- data.frame(X1, X2, W, Y)
head(data)
```

---

## 2. 🎯 Estimate Propensity Scores

```{r}
ps_model <- glm(W ~ X1 + X2, data = data, family = binomial())
data$pscore <- predict(ps_model, type = "response")
```

---

## 3. ⚖️ Compute Weights (Inverse Probability of Treatment Weights)

```{r}
data$wt <- ifelse(data$W == 1, 1 / data$pscore, 1 / (1 - data$pscore))
summary(data$wt)
```

---

## 4. 🧮 Estimate Weighted Treatment Effect

```{r}
design <- svydesign(ids = ~1, weights = ~wt, data = data)
model_psw <- svyglm(Y ~ W, design = design)
summary(model_psw)
```

---

## 5. 📊 Visualize Propensity Score Overlap

```{r}
ggplot(data, aes(x = pscore, fill = factor(W))) +
  geom_density(alpha = 0.5) +
  labs(title = "Propensity Score Distribution by Treatment",
       x = "Propensity Score", fill = "Treatment") +
  theme_minimal()
```

---

## ✅ Summary

PSW reduces confounding by **balancing covariates** across treatment groups, allowing us to estimate causal effects in non-randomized data.

---

## 📖 References

- Rosenbaum & Rubin (1983). *The Central Role of the Propensity Score in Observational Studies*
- Stuart (2010). *Matching Methods for Causal Inference: A Review and a Look Forward*