---
title: "Sensitivity Analysis: How Robust Is Your Causal Inference?"
output: rmarkdown::html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(dplyr)
library(ggplot2)
# Optional: Uncomment to use sensemakr
# install.packages("sensemakr")
# library(sensemakr)
```

## ðŸŽ¯ Sensitivity Analysis: Why It Matters

Even when using careful matching or regression, unobserved confounding may bias our treatment effect estimates. Sensitivity analysis helps us **quantify how much unobserved bias would be required** to nullify our results.

---

## 1. ðŸ§ª Simulate Confounded Data

```{r}
set.seed(123)
n <- 1000

U <- rnorm(n)  # unobserved confounder
X <- rbinom(n, 1, prob = plogis(0.3 * U))  # treatment depends on U
Y <- 2 * X + 1.5 * U + rnorm(n)  # outcome depends on X and U

data <- data.frame(X, Y, U)
```

---

## 2. âš ï¸ Naive Regression (No U Adjustment)

```{r}
naive_model <- lm(Y ~ X, data = data)
summary(naive_model)
```

---

## 3. ðŸ“ True Model with U (for comparison)

```{r}
true_model <- lm(Y ~ X + U, data = data)
summary(true_model)
```

---

## 4. ðŸ“‰ Sensitivity with Manual Bias Factor

We'll manually simulate how adding different levels of confounding changes the treatment effect:

```{r}
bias_sim <- data.frame(
  rho = seq(0, 1, by = 0.05)
) %>%
  mutate(
    bias = rho * sd(U),
    adjusted_beta = coef(naive_model)[["X"]] - bias
  )

ggplot(bias_sim, aes(x = rho, y = adjusted_beta))
  geom_line(color = "steelblue", size = 1)
  geom_hline(yintercept = coef(true_model)[["X"]], linetype = "dashed", color = "red")
  labs(title = "Effect of Unobserved Confounding on Estimated Treatment Effect",
       x = "Assumed Correlation with U (rho)", y = "Adjusted Beta for X")
  theme_minimal()
```

---

## âœ… Summary

Sensitivity analysis gives us a **range of possible estimates** under different assumptions about unobserved confounders. It helps answer:

> "How strong would the unobserved confounding need to be to explain away my effect?"

---

## ðŸ“– References

- Cinelli & Hazlett (2020). *Making sense of sensitivity: Extending omitted variable bias.*
- Rosenbaum (2002). *Observational Studies.*