---
title: "Synthetic Control Methods for Policy Evaluation"
output: rmarkdown::html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(dplyr)
library(ggplot2)
# Optional: install.packages("Synth")
```

## 🧪 Synthetic Control Methods

Synthetic Control Methods (SCM) estimate the effect of an intervention by comparing the **treated unit** to a **weighted combination of control units** that best resemble it in the pre-treatment period.

---

## 1. ✨ Simulate Panel Data

```{r}
set.seed(123)
n_units <- 10
n_years <- 20

unit <- rep(1:n_units, each = n_years)
year <- rep(2000:2019, times = n_units)
treated <- ifelse(unit == 1 & year >= 2010, 1, 0)  # treatment begins in 2010 for unit 1

# Common trend + unit effects
alpha <- rnorm(n_units, 0, 1)
trend <- 0.5 * (year - 2000)
error <- rnorm(n_units * n_years)

Y <- 5 + alpha[unit] + trend + treated * 3 + error

df <- data.frame(unit = as.factor(unit), year, treated, Y)
head(df)
```

---

## 2. 📈 Plot Treated vs. Controls Over Time

```{r}
ggplot(df, aes(x = year, y = Y, group = unit, color = unit == 1))
  geom_line()
  labs(title = "Treated Unit vs Controls Over Time", color = "Treated")
  theme_minimal()
```

---

## 3. 🧮 Synthetic Control with Pre-treatment Matching

Here’s a simplified SCM approach: match pre-treatment means.

```{r}
df_pre <- df %>% filter(year < 2010)
df_post <- df %>% filter(year >= 2010)

avg_pre <- df_pre %>% group_by(unit) %>% summarise(mean_pre = mean(Y))
weights <- avg_pre %>% filter(unit != 1) %>%
  mutate(weight = 1 / abs(mean_pre - avg_pre$mean_pre[1])) %>%
  mutate(weight = weight / sum(weight))

synth_post <- df_post %>%
  filter(unit != 1) %>%
  left_join(weights, by = "unit") %>%
  group_by(year) %>%
  summarise(synth_Y = sum(Y * weight))

actual_post <- df_post %>% filter(unit == 1) %>% select(year, Y)

synth_df <- left_join(actual_post, synth_post, by = "year")
```

---

## 4. 📊 Plot Synthetic vs. Treated

```{r}
ggplot(synth_df, aes(x = year))
  geom_line(aes(y = Y), color = "blue", size = 1.2)
  geom_line(aes(y = synth_Y), color = "orange", linetype = "dashed", size = 1.2)
  labs(title = "Synthetic Control vs Treated Unit (Post-2010)", y = "Outcome", x = "Year")
  theme_minimal()
```

---

## ✅ Summary

Synthetic Control Methods:
- Create a data-driven counterfactual
- Are especially useful for **policy evaluation**
- Can be extended using `Synth`, `tidysynth`, or Bayesian SCM tools

---

## 📖 References

- Abadie, Diamond & Hainmueller (2010). *Synthetic Control Methods for Comparative Case Studies*
- Abadie (2021). *Using Synthetic Controls for Causal Inference*